On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-05-11 Manuscript/manuscript.tex
Starting to work again

\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
clustered data with subunit-level grouping. The test is designed based on a
within-cluster resampling principle. Consider randomly picking one observation
from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
Wilcoxon rank-sum statistic for the pseudo-sample is
\[
W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
\]
where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
The test statistic $S$ is the average of $W^*$ averaged over all possible
pseudo-samples conditioning on the observed data and group indicators.
The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
standardized to form a statistic $Z$ which follows standard normal distribution
asymptotically \citep[p.910]{datta2005rank}. An implementation of this method is
available from the \texttt{clusWilcox.test()} function with
\texttt{method = ``ds'} from R package \texttt{clusrank}
\citep{jiang2017wilcoxon}.
\eds{are you including preliminary heat data as possible within cluster 
observations in this analysis?  Based on your 
discussion in Section 2.1 about how they are not comparable, it may not be fair 
to include them here?  Or perhaps do the analysis with and without prelim heat
data to see if it makes a difference?}

@@

Devon Allen's disqualification at the World Track and Field Championships was
possibly the result of both a faulty timing equipment and Devon Allen
reacting extremely quickly to the start.  Our analysis showed that
the 2022 World Track and Field Championships were a low outlier compared to
every other year in terms of average reaction time.  Additionally, it was shown
through a gamma mixed effects model that there are significant year and race
effects \eds{remove race effects?} that greatly impact reaction time.  
However, this practically does not make much sense
as there is no reason for so much random variation without much of a trend.
There has not been a consistent decrease or increase since 1999, and much of the 
data
for reaction times has been random and unpredictable. 
\eds{I would suggest removing the last two sentences, as they would suggest your
 modeling with GLMM is not good}.

This paper does not consider data from the women's 100 meter hurdle, and although
that was initially because of concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2022 
\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
the intro, unless he only considered sprint and not hurdles?} Likewise,
the rank-based comparisons resulted in p values that were not significant.  The
lack of evidence from the women's data makes it harder to conclude 
that the sensors used at the 2022 championships were faulty, although it is
unknown if the sensors for the men's 110 meter hurdle and women's 100 meter
hurdle were the same.

>>>>

\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
clustered data with subunit-level grouping. The test is designed based on a
within-cluster resampling principle. Consider randomly picking one observation
from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
Wilcoxon rank-sum statistic for the pseudo-sample is
\[
W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
\]
where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
The test statistic $S$ is the average of $W^*$ averaged over all possible
pseudo-samples conditioning on the observed data and group indicators.
The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
standardized to form a statistic $Z$ which follows standard normal distribution
asymptotically \citep[p.910]{datta2005rank}. An implementation of this method is
available from the \texttt{clusWilcox.test()} function with
\texttt{method = `ds'} from R package \texttt{clusrank}
\citep{jiang2017wilcoxon}.
\eds{are you including preliminary heat data as possible within cluster 
observations in this analysis?  Based on your 
discussion in Section 2.1 about how they are not comparable, it may not be fair 
to include them here?  Or perhaps do the analysis with and without prelim heat
data to see if it makes a difference?}

@@

Devon Allen's disqualification at the World Track and Field Championships was
possibly the result of both a faulty timing equipment and Devon Allen
reacting extremely quickly to the start.  Our analysis showed that
the 2022 World Track and Field Championships were a low outlier compared to
every other year in terms of average reaction time.  Additionally, it was shown
through a gamma mixed effects model that there are significant year and race
effects \eds{remove race effects?} that greatly impact reaction time.

\eds{I would suggest removing the last two sentences, as they would suggest your
 modeling with GLMM is not good}.  
\of{I agree and they have been removed}

This paper does not consider data from the women's 100 meter hurdle, and although
that was initially because of concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2003 
\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
the intro, unless he only considered sprint and not hurdles?} Likewise,
the rank-based comparisons resulted in p values that were not significant.  The
lack of evidence from the women's data makes it harder to conclude 
that the sensors used at the 2022 championships were faulty, although it is
unknown if the sensors for the men's 110 meter hurdle and women's 100 meter
hurdle were the same.
